Multi-wavelength Radio Continuum Emission Studies of Dust-free Red Giants

### 2013-04-01 draft_06/draft.tex
added draft_06

Multi-wavelength centimeter continuum observations of non-dusty red giants directly sample their wind acceleration zones where most of the energy that drives their mass loss is being deposited. These stars are feeble thermal emitters at these wavelengths however, and previous observations have provided only a small number of modest signal-to-noise measurements accumulated over many years. Here, we present multi-wavelength Karl G. Jansky Very Large Array (VLA) thermal continuum observations of the acceleration zones of two dust-free red giants, Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). Uniquely, these multi-wavelength observations were mostly carried out within a few days of each other so any long term variability that may exist can be ignored. We report the first detections at many wavelengths for each star including a detection at 10 cm (3.0 GHz; S-band) for both stars and a 20 cm (1.5 GHz; L-band) detection for $\alpha$ Boo. This is the first time single (non-binary) luminosity class III red giants have been observed at such long continuum wavelengths. These long wavelength flux density measurements sample the outer layers of the stars atmosphere where the wind velocity is approaching (or possibly has reached) its terminal value and the ionization balance is becoming \textit{frozen-in}. As a result, they are important empirical inputs for future atmospheric models. We present spectral energy distributions for both stars and compare these with previous measurements and published semi-empirical models. The marked deviations between our data and existing semi-empirical models highlight the need for new atmospheric models to be developed. The derived spectral indices are discussed in relation to the possible properties of the stellar atmospheres and a simple analytical atmospheric model is developed for $\alpha$ Boo which is in close agreement to the observed flux densities at long wavelengths.

@@

Mass loss from late-type evolved stars plays a crucial role in both stellar and galactic evolution and ultimately provides part of of the material required for the next generation of stars and planets. Despite the importance of this phenomenon and decades of study, the mechanisms that drive winds from evolved spectral-type K through mid-M stars remain an enduring mystery (clearly laid out by \cite{1985ASSL..117..229H} but still unsolved, e.g.
\citealt{2009AIPC.1094..267C}). There is insufficient atomic, molecular, or dust opacity to drive a radiation-driven outflow and acoustic/pulsation models cannot drive the observed mass loss rates \citep{1995ApJ...442L..61S}. Optical and ultraviolet observations reveal an absence of hot wind plasma and the winds are thus too cool to be Parker-type thermally-driven flows \cite[e.g.][]{1979ApJ...229L..27L,1981ApJ...250..293A}.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function is very sensitive to electron temperature (i.e. $S_{\nu} \propto e^{-h\nu /kT}$) and so a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and not reflect the global mean value. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e. $S_{\nu} = {2kT\nu ^2 /c^2}$) and so should give a more appropriate estimate of the thermal radial wind structure. It is this value that controls the atomic level populations and ionization that feed into UV spectroscopic analyses. This value is also needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, and thus derive constraints on potential mass loss mechanisms.

In the cm-radio regime the radio opacity strongly increases with wavelength (i.e. $ \kappa \propto \lambda ^{2.1}$) so the longer wavelengths sample the extended layers of the stars atmosphere thus essentially providing us with spatial information about the star's mass outflow region. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at these radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new Karl G. Jansky Very Large Array (VLA) observations with atmospheric models derived from UV analysis.

@@

Currently the most detailed spatial information about the atmospheres of K and early M evolved stars is obtained from eclipsing binaries such as the symbiotic and $\zeta$ Aurigae systems (e.g. \citealt{1996ApJ...466..979B}; \citealt{2008AJ....136.1964E}; \citealt{2008ApJ...675..711C}). Even though these systems offer us the best opportunity to obtain information on the dynamics and thermodynamics at various heights in the evolved star's atmosphere, the very nature of the binary system may introduce further complexities. For example, the orbital separation is often within the wind acceleration region and one could expect wind accretion and flow perturbations to be present (e.g. \citealt{1981ApJ...248.1043C}). In fact, using the `old' VLA, \cite{2005AJ....129.1018H} confirm that the velocity structure of  of $\zeta$ Aurigae is not typical of single stars with similar spectral types.

In order to avoid the assumed additional complexities of a companion, we have selected two single luminosity class III red giants: Arcturus ($\alpha$ Boo: K2 III) and Aldebaren ($\alpha$ Tau: K5 III). These nearby red giants have been extensively studied at other wavelengths and their stellar parameters, which are briefly summarized in Table 1, are accurately known. These stars are predicted to be point sources at all frequencies in all VLA configurations so our radio observations measure their total flux density, $F_{\nu}$. Moreover both stars have existing semi-empirical 1-D chromospheric and wind models which we directly compare to our data in this paper.

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablecaption{Some properties of our red giant sample.}
\tablehead{	\colhead{}		       				& 
			\colhead{$\alpha$ Boo}				&
			\colhead{$\alpha$ Tau}				}
\startdata
Spectral Type 				& K2 III  & K5 III  \\
Effective Temperature (K)	& $4294 \pm 30$  & $3970 \pm 49$ \\
Distance (pc)	& 20.43  & 11.26 \\
Angular Diameter (mas)		& $21.0 \pm 0.2$ & $20.2 \pm 0.3$ \\
Mass Loss Rate (M$_{\odot}$ yr$^{-1}$)		& 2$\times$10$^{-10}$ & 1.6$\times$10$^{-11}$ \\
Wind Temperature (K)		& $\sim 10,000$  & $< 10,000$  \\
Wind Terminal Velocity & $\sim$ 40 km s$^{-1}$ & $\sim$ 30 km s$^{-1}$ \\
Semi-empirical model	& \cite{1985pssl.proc..351D} & \cite{1999MNRAS.302...37M}
\enddata
\tablecomments{Effective temperatures and photospheric diameters are from \cite{1993AA...270..315D}. Distances are from \cite{2007A&A...474..653V}. Mass loss rates are from \cite{1985pssl.proc..351D} and \cite{1998ApJ...503..396R}.}

@@

Both $\alpha$ Boo and $\alpha$ Tau were slightly offset from the phase-center by $\sim$\,5 synthesized beam widths in order to avoid possible errors at phase-center. All scheduling blocks (SBs) were kept to $\le$\,2.5 hours of duration to improve their likelihood of being scheduled. For the high frequency observations (i.e. Q, Ka, and K-bands) we used the \textit{fast switching} technique which consisted of rapidly alternating observations of the target source and a nearby unresolved phase calibrator. The total cycle times for the Q, Ka, and K-band observations were 160, 230, and 290 s, respectively. For both target sources these high frequency observations were combined into a single 2 hour observing track and commenced with X-band reference pointing with solutions being applied on-line. After X-band pointing the target source was observed at Q-band to insure the best pointing solutions were used. The lower frequencies tracks were composed of repeatedly interleaving observations of the target source and a nearby phase calibrator but had much longer cycle times. The primary calibration sources 3C286 and 3C138 were observed at the end of all tracks and were used to measure the complex bandpass and set the absolute flux for $\alpha$ Boo and $\alpha$ Tau, respectively.

The data were flagged, calibrated, and imaged within the Common Astronomical Software Application  \cite[CASA;][]{2007ASPC..376..127M} package. Data deemed to be bad by the VLA online system were flagged, as were pure zeros, non-operational antennae, dummy scans at the beginning of each track, and poorly performing antennae. Visual inspection of each scan was carried out to determine if data at the beginning or end of these scans needed to be flagged (a process known as \textit{quacking}). For the 2011 low frequency data the two sub-bands were centered at relatively radio frequency interference (RFI) free regions of the bandpass and only a very small amount of RFI had to be flagged.   The 2012 wide-band data was initially Hanning smoothed (combining adjacent frequency channels with weights 0.25, 0.5, and 0.25) to suppress Gibbs ringing. We manually flagged entire sub-bands that were badly contaminated with RFI. The \textit{testautoflag} task was then used to conservatively flag RFI from all sources and any remaining RFI was manually flagged.

@@

Three different methods were used to calculate the flux density from the unresolved target sources in each image:
\begin{enumerate}
\item by taking the peak pixel value from the source
\item by manually integrating the flux density around the source
\item by fitting an elliptical Gaussian model to the source and deriving the integrated flux density using the CASA \textit{imfit} task.
\end{enumerate}
Each of these values along with the image root mean square (rms) noise measured from adjacent background regions and fitting error produced by \textit{imfit} are given in Table \ref{tab:tab3}. We assume absolute flux density scale systematic uncertainties of  3\% at all frequencies.

\begin{deluxetable*}{Cccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Fluxe Densities of $\alpha$ Boo and $\alpha$ Tau}
\tablehead{\colhead{}		            &
			\colhead{Band}		            &
			\colhead{Frequency\tablenotemark{a}}&
			\colhead{Peak $F_{\nu}$}					&
          	\colhead{Integrated $F_{\nu}$} 			&
          	\colhead{\textit{Imfit} Integrated}	&
			\colhead{Image rms}					&
           	\colhead{\textit{Imfit} Fitting}  	\\
	\colhead{}									&
	\colhead{}									&
	\colhead{(GHz)}								&
    \colhead{(mJy)}		        	& 
    \colhead{(mJy)}		                		& 
	\colhead{$F_{\nu}$ (\rm{mJy})}                        		& 
	\colhead{(Jy beam^{-1})}         					&
	\colhead{Error (mJy)}		}
\startdata
$\alpha$ Boo (K2 III) &Q  &43.28& 5.941 & 6.093 & 6.420 & 0.301 &  0.261\\
&Ka &33.56& 4.159 & 4.319 & 4.489 & 0.083 & 0.090 \\
&K  &22.46& 1.827 & 1.784 & 1.809 & 0.043 & 0.050 \\
&X  &8.46 & 0.510 & 0.514 & 0.532 & 0.030 & 0.019 \\
&C  &4.90 & 0.214 & 0.144 & 0.159 & 0.035 & 0.009 \\
&S  &3.15 & 0.148 & 0.135 & 0.160 & 0.028 & 0.010\\
&S  &2.87 & 0.127 & 0.118 & 0.116 & 0.012 & 0.016\\
&L  &1.63 & 0.067 & 0.068 & 0.090 & 0.013 & 0.015\\
\hline
\rule{0pt}{3ex}  $\alpha$ Tau (K5 III)&Q  &43.28 & 3.672& 3.734 & 4.082 &  0.259& 0.183	\\
&Ka &33.56 & 2.188& 1.962 & 2.125 &  0.091& 0.070 \\
&K  &22.46 & 1.864& 1.881 & 2.069 &  0.042& 0.083 \\
&X  &8.46  & 0.296& 0.287 & 0.281 &  0.014& 0.015 \\
&C  &4.96  & 0.147& 0.167 & 0.176 &  0.010& 0.010 \\
&S  &3.15  & 0.062& 0.043 & - &  0.017 & -
\enddata
\tablenotetext{a}{Frequency of final image produced using \textit{clean}'s multi-frequency synthesis imaging mode.}
\label{tab:tab3}
\end{deluxetable*}

@@

Apart from $\alpha$ Boo at C-band and $\alpha$ Tau at S-band, detections were made in both sub-bands for the 2011 data. The flux densities of the targets in each sub-bands were found to be the same within their uncertainties so we do not present these values here. Instead we give the values from the radio maps produced by concatenating the two sub-bands. We present in Table \ref{tab:tab3} the target flux densities extracted from these concatenated radio maps. In the following two sections we briefly discuss the properties of these radio maps for both targets.

\subsection{$\alpha$ Boo Radio Maps} \label{results1} 
High S/N detections ($>19\sigma$) of $\alpha$ Boo were made at 22.5, 33.6, and 43.3 GHz. Some residuals of the dirty beam remained in the CLEANed maps due to the paucity of uv-coverage in these short high frequency observations. The lower frequencies maps were contaminated by the emission of a strong radio source located $186\arcsec$ north-west of $\alpha$ Boo. This source was reported by \cite{1986AJ.....91..602D} and their flux of 25 mJy at 4.9 GHz is in close agreement with our measurement of 23.2 mJy at the same frequency. We find the source to have a spectral index of -1.4 and its flux reaches  80.3 mJy at 1.6 GHz. Surprisingly we do not detect $\alpha$ Boo in the higher frequency sub-band at 5.0 GHz (C-band) so the values given in Table \ref{tab:tab3} are taken from the lower frequency sub-band only. We obtain good detections ($>5\sigma$) of the star for both epochs at $\sim$3 GHz (S-band) and the peak flux densities agree within their uncertainties. We can therefore safely assume that the 1.5 GHz (L-band) flux has not changed significantly over that period either, and so can safely be included in any analysis. The map at L-band was highly contaminated by the emission of the strong source north-west of $\alpha$ Boo but the star is still detected at the $5\sigma$ level. There is a slight positional offset of $1\arcsec$ between the position of the peak flux density at 1.5 and at 3.0 GHz for the 2012 data, which were taken within 1 day of each other. However, this is less than a quarter of the 1.5 GHz synthesized beam and considering that the $4\sigma$ contours overlap, we feel that it is highly unlikely that both sources are not $\alpha$ Boo.

\subsection{$\alpha$ Tau Radio Maps} \label{results2}
The final deconvolved radio maps of $\alpha$ Tau were of excellent quality with the rms noise reaching the predicted thermal noise in many cases. The target field at all frequencies was free from strong serendipitous radio sources and thus the final images were free of the sidelobe contamination that were present in the low frequency $\alpha$ Boo images. $\alpha$ Tau was the only source in the high frequency maps while the brightest source in the low frequency maps was located $106\arcsec$ north north-east of $\alpha$ Tau and had flux densities of 0.85, 1.35, and 1.7 mJy at 8.5, 5.0, and 3.5 GHz, respectively. Strong detections ($>14\sigma$) of $\alpha$ Tau were made at all frequencies between 5.0 and 43.3 GHz. Due to the limited number of S-band receivers installed at the time, a full 2.5 hr track was dedicated to $\alpha$ Tau at 3.1 GHz in order to achieve the required sensitivity to give a possible detection. We report a tentative $3\sigma$ detection of $\alpha$ Tau at 3.1 GHz when we take its peak pixel value as its total flux density. As this is a weak detection we avoid using the \textit{imfit} task to obtain a flux density estimate, as this may introduce further errors \citep{1999ASPC..180.....T}.

>>>>

Multi-wavelength centimeter continuum observations of non-dusty red giants directly sample their wind acceleration zones where most of the energy that drives their mass loss is being deposited. These stars are feeble thermal emitters at these wavelengths however, and previous observations have provided only a small number of modest signal-to-noise measurements accumulated over many years. Here, we present multi-wavelength Karl G. Jansky Very Large Array (VLA) thermal continuum observations of the acceleration zones of two dust-free red giants, Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). Uniquely, our observations were mostly carried out within a few days of each other so any long term variability that may exist can be ignored. We report the first detections at many wavelengths for each star including a detection at 10 cm (3.0 GHz; S-band) for both stars and a 20 cm (1.5 GHz; L-band) detection for $\alpha$ Boo. This is the first time single (non-binary) luminosity class III red giants have been observed at such long continuum wavelengths. These long wavelength flux density measurements sample the outer layers of the stars atmosphere where the wind velocity is approaching (or possibly has reached) its terminal value and the ionization balance is becoming \textit{frozen-in}. As a result, they are important empirical inputs for future atmospheric models. We present spectral energy distributions for both stars and compare these with previous measurements and published semi-empirical models based on ultraviolet data. The marked deviations between our data and existing semi-empirical models highlight the need for new atmospheric models to be developed. The derived spectral indices are discussed in relation to the possible properties of the stellar atmospheres and a simple analytical atmospheric model is developed for $\alpha$ Boo's outer atmosphere which is in good agreement with our new long wavelength flux densities.

@@

Mass loss from late-type evolved stars plays a crucial role in both stellar and galactic evolution and ultimately provides part of of the material required for the next generation of stars and planets. Despite the importance of this phenomenon and decades of study, the mechanisms that drive winds from evolved spectral-type K through mid-M stars remain an enduring mystery [clearly laid out by \cite{1985ASSL..117..229H} but still unsolved, e.g.,
\citealt{2009AIPC.1094..267C}]. There is insufficient atomic, molecular, or dust opacity to drive a radiation-driven outflow and acoustic/pulsation models cannot drive the observed mass loss rates \citep{1995ApJ...442L..61S}. Optical and ultraviolet observations reveal an absence of hot wind plasma and the winds are thus too cool to be Parker-type thermally-driven flows \cite[e.g.,][]{1979ApJ...229L..27L,1981ApJ...250..293A}.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function is very sensitive to electron temperature (i.e., $S_{\nu} \propto e^{-h\nu /kT}$) and so a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and not reflect the global mean value. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e., $S_{\nu} = {2kT\nu ^2 /c^2}$) and so should give a more appropriate estimate of the thermal radial wind structure. It is this value that controls the atomic level populations and ionization that feed into UV spectroscopic analyses. This value is also needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, and thus derive constraints on potential mass loss mechanisms.

In the cm-radio regime the radio opacity strongly increases with wavelength (i.e., $ \kappa _{\lambda} \propto \lambda ^{2.1}$) and so the longer wavelengths sample the extended layers of the stars atmosphere, thus essentially providing us with spatial information about the star's mass outflow region. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at these radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new NRAO\footnote{The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} Karl G. Jansky Very Large Array (VLA) observations with atmospheric models derived from UV analysis.

@@

Currently the most detailed spatial information about the atmospheres of K and early M evolved stars is obtained from eclipsing binaries such as the symbiotic and $\zeta$ Aurigae systems (e.g., \citealt{1996ApJ...466..979B}; \citealt{2008AJ....136.1964E}; \citealt{2008ApJ...675..711C}). Even though these systems offer us the best opportunity to obtain information on the dynamics and thermodynamics at various heights in the evolved star's atmosphere, the very nature of the binary system may introduce further complexities. For example, the orbital separation is often within the wind acceleration region and one could expect flow perturbations to be present (e.g., \citealt{1981ApJ...248.1043C}). In fact, using the `old' VLA, \cite{2005AJ....129.1018H} confirm that the velocity structure of  of $\zeta$ Aurigae is not typical of single stars with similar spectral types, such as $\lambda$ Velorum \citep{1999ApJ...521..382C}.

In order to avoid the assumed additional complexities of a companion, we have selected two single luminosity class III red giants: Arcturus ($\alpha$ Boo: K2 III) and Aldebaren ($\alpha$ Tau: K5 III). These nearby red giants have been extensively studied at other wavelengths and their stellar parameters, which are briefly summarized in Table 1, are accurately known. These stars are predicted to be point sources at all frequencies in all VLA configurations so our radio observations measure their total flux density, $F_{\nu}$. Moreover, both stars have existing semi-empirical 1-D chromospheric and wind models which we directly compare to our data in this paper.

\begin{deluxetable}{ccc}
\tabletypesize{\scriptsize}
\tablecaption{Some properties of our red giant sample.}
\tablehead{	\colhead{}		       				& 
			\colhead{$\alpha$ Boo}				&
			\colhead{$\alpha$ Tau}				}
\startdata
Spectral Type 				& K2 III  & K5 III  \\
Effective Temperature (K)	& $4294 \pm 30$  & $3970 \pm 49$ \\
Angular Diameter (mas)		& $21.0 \pm 0.2$ & $20.2 \pm 0.3$ \\
Distance (pc)	& $11.3 \pm 0.1$ & $20.4 \pm 0.3$\\
Radius ($R_{\odot}$)	& $25.4 \pm 0.3$  & $44.4 \pm 1.0$ \\
Mass Loss Rate ($M_{\odot}$ yr$^{-1}$)		& 2$\times$10$^{-10}$ & 1.6$\times$10$^{-11}$ \\
Wind Temperature (K)		& $\sim10,000$  & $< 10,000$  \\
Wind Terminal Velocity & $\sim$40 km s$^{-1}$ & $\sim$30 km s$^{-1}$ \\
Semi-empirical model	& \cite{1985pssl.proc..351D} & \cite{1999MNRAS.302...37M}
\enddata
\tablecomments{Effective temperatures and photospheric angular diameters are from \cite{1993AA...270..315D}. Distances are from \cite{2007A&A...474..653V}. Mass loss rates are from \cite{1985pssl.proc..351D} and \cite{1998ApJ...503..396R}.}

@@

Both $\alpha$ Boo and $\alpha$ Tau were slightly offset from the phase-center by $\sim$\,5 synthesized beam widths in order to avoid possible errors at phase-center. All scheduling blocks (SBs) were kept to $\le$\,2.5 hours of duration to improve their likelihood of being scheduled. For the high frequency observations (i.e., Q, Ka, and K-bands) we used the \textit{fast switching} technique which consisted of rapidly alternating observations of the target source and a nearby unresolved phase calibrator. The total cycle times for the Q, Ka, and K-band observations were 160, 230, and 290 s, respectively. For both target sources these high frequency observations were combined into a single 2 hour observing track and commenced with X-band reference pointing with solutions being applied on-line. After X-band pointing the target source was observed at Q-band to insure the best pointing solutions were used. The lower frequencies tracks were composed of repeatedly interleaving observations of the target source and a nearby phase calibrator but had much longer cycle times. The primary calibration sources 3C286 and 3C138 were observed at the end of all tracks and were used to measure the complex bandpass and set the absolute flux for $\alpha$ Boo and $\alpha$ Tau, respectively.

The data were flagged, calibrated, and imaged within the Common Astronomical Software Application  \cite[CASA;][]{2007ASPC..376..127M} package. Data deemed to be bad by the VLA online system were flagged, as were pure zeros, non-operational antennae, dummy scans at the beginning of each track, and poorly performing antennae. Visual inspection of each scan was carried out to determine if data at the beginning or end of these scans needed to be flagged. For the 2011 low frequency data the two sub-bands were centered at relatively radio frequency interference (RFI) free regions of the bandpass and only a very small amount of RFI had to be flagged.   The 2012 wide-band data were initially Hanning smoothed (combining adjacent frequency channels with weights 0.25, 0.5, and 0.25) to suppress Gibbs ringing. We manually flagged entire sub-bands that were badly contaminated with RFI. The \textit{testautoflag} task was then used to conservatively flag RFI from all sources and any remaining RFI was manually flagged.

@@

In each image the flux density from the unresolved target source was calculated by, 1) taking the peak pixel value from the source, 2) manually integrating the flux density around the source, and 3) fitting an elliptical Gaussian model to the source and deriving the integrated flux density using the CASA \textit{imfit} task. Each of these values along with the image root mean square (rms) noise measured from adjacent background regions and fitting error produced by \textit{imfit} are given in Table \ref{tab:tab3}. We assume absolute flux density scale systematic uncertainties of  3\% at all frequencies \citep{2013ApJS..204...19P}.

\begin{deluxetable*}{Cccccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{VLA Fluxe Densities of $\alpha$ Boo and $\alpha$ Tau}
\tablehead{\colhead{}		            &
			\colhead{Band}		            &
			\colhead{Frequency\tablenotemark{a}}&
			\colhead{Peak $F_{\nu}$}					&
          	\colhead{Integrated $F_{\nu}$} 			&
          	\colhead{\textit{Imfit} Integrated}	&
			\colhead{Image rms}					&
           	\colhead{\textit{Imfit} Fitting}  	\\
	\colhead{}									&
	\colhead{}									&
	\colhead{(GHz)}								&
    \colhead{(mJy)}		        	& 
    \colhead{(mJy)}		                		& 
	\colhead{$F_{\nu}$ (\rm{mJy})}                        		& 
	\colhead{(Jy beam^{-1})}         					&
	\colhead{Error (mJy)}		}
\startdata
$\alpha$ Boo (K2 III) &Q  &43.28& 5.941 & 6.093 & 6.420 & 0.301 &  0.261\\
&Ka &33.56& 4.159 & 4.319 & 4.489 & 0.083 & 0.090 \\
&K  &22.46& 1.827 & 1.784 & 1.809 & 0.043 & 0.050 \\
&X  &8.46 & 0.510 & 0.514 & 0.532 & 0.030 & 0.019 \\
&C  &4.90 & 0.214 & 0.144 & 0.159 & 0.035 & 0.009 \\
&S  &3.15 & 0.148 & 0.135 & 0.160 & 0.028 & 0.010\\
&S  &2.87 & 0.127 & 0.118 & 0.116 & 0.012 & 0.016\\
&L  &1.63 & 0.067 & 0.068 & 0.090 & 0.013 & 0.015\\
\hline
\rule{0pt}{3ex}  $\alpha$ Tau (K5 III)&Q  &43.28 & 3.672& 3.734 & 4.082 &  0.259& 0.183	\\
&Ka &33.56 & 2.188& 1.962 & 2.125 &  0.091& 0.070 \\
&K  &22.46 & 1.864& 1.881 & 2.069 &  0.042& 0.083 \\
&X  &8.46  & 0.296& 0.287 & 0.281 &  0.014& 0.015 \\
&C  &4.96  & 0.147& 0.167 & 0.176 &  0.010& 0.010 \\
&S  &3.15  & 0.062& 0.043 & - &  0.017 & -
\enddata
\tablenotetext{a}{Frequency of the final image produced using the multi-frequency synthesis imaging mode within CASA's \textit{clean} task.}
\label{tab:tab3}
\end{deluxetable*}

@@

Apart from $\alpha$ Boo at C-band and $\alpha$ Tau at S-band, detections were made in both sub-bands for the 2011 data. At each other frequency, the flux densities of the targets in both sub-bands were found to be the same within their uncertainties so we do not present these values here. Instead we give the values from the radio maps produced by concatenating the two sub-bands. We present in Table \ref{tab:tab3} the target flux densities extracted from these concatenated radio maps. In the following two sections we briefly discuss the properties of these radio maps for both targets.

\subsection{$\alpha$ Boo Radio Maps} \label{results1} 
High S/N detections ($>19\sigma$) of $\alpha$ Boo were made at 22.5, 33.6, and 43.3 GHz. Some residuals of the dirty beam remained in the CLEANed maps due to the paucity of uv-coverage in these short high frequency observations. The lower frequencies maps were contaminated by the emission of a strong radio source located $186\arcsec$ north-west of $\alpha$ Boo. This source was reported by \cite{1986AJ.....91..602D} and their flux density of 25 mJy at 4.9 GHz is in close agreement with our measurement of 23.2 mJy at the same frequency. We find the source to have a spectral index of -1.4 and reaches 80.3 mJy at 1.6 GHz. Surprisingly we do not detect $\alpha$ Boo in the higher frequency sub-band at 5.0 GHz (C-band) so the values given in Table \ref{tab:tab3} are taken from the lower frequency sub-band only. We obtain good detections ($>5\sigma$) of the star for both epochs at $\sim$3 GHz (S-band) and the peak flux densities agree within their uncertainties. We can therefore safely assume that the 1.5 GHz (L-band) flux density has not changed significantly over that period either, and so can safely be included in any analysis. The map at L-band was highly contaminated by the sidelobes of the strong source north-west of $\alpha$ Boo but the star is still detected at the $5\sigma$ level. There is a slight positional offset of $1\arcsec$ between the position of the peak flux density at 1.5 and at 3.0 GHz for the 2012 data, which were taken within 1 day of each other. However, this is less than a quarter of the 1.5 GHz synthesized beam and considering that the $4\sigma$ contours overlap, we feel that it is highly unlikely that both sources are not $\alpha$ Boo.

\subsection{$\alpha$ Tau Radio Maps} \label{results2}
The final deconvolved radio maps of $\alpha$ Tau were of excellent quality with the rms noise reaching the predicted thermal noise in many cases. The target field at all frequencies was free from strong serendipitous radio sources and thus the final images were free of the sidelobe contamination that were present in the low frequency $\alpha$ Boo images. $\alpha$ Tau was the only source in the high frequency maps while the brightest source in the low frequency maps was located $106\arcsec$ north north-east of $\alpha$ Tau and had flux densities of 0.85, 1.35, and 1.7 mJy at 8.5, 5.0, and 3.5 GHz, respectively. Strong detections ($>14\sigma$) of $\alpha$ Tau were made at all frequencies between 5.0 and 43.3 GHz. Due to the limited number of S-band receivers installed at the time, a full 2.5 hr track was dedicated to $\alpha$ Tau at 3.1 GHz in order to achieve the required sensitivity to give a possible detection. We report a tentative $3\sigma$ detection of $\alpha$ Tau at 3.1 GHz when we take its peak pixel value as its total flux density. As this is a weak detection we avoid using the \textit{imfit} task to obtain a flux density estimate, as this may introduce additional errors \citep{1999ASPC..180.....T}.

\begin{deluxetable*}{Ccccc}
\tabletypesize{\scriptsize}
%\tablecolumns{6} 
%\tablewidth{0pt} 
\tablecaption{Compilation of Previous Radio Observations of $\alpha$ Boo and $\alpha$ Tau ($\nu \le 250$ GHz)}
\tablehead{							            &
    \colhead{Frequency}		       				& 
			\colhead{Date}						&
			\colhead{Flux Density}				&
          	\colhead{Source} 			\\
												&
	\colhead{(GHz)}								&
	\colhead{}									&
    \colhead{(mJy)}		        				& 
    \colhead{}			         				}
\startdata
$\alpha$ Boo (K2 III) &4.9  & 1983 Jan 21 & $0.39 \pm 0.13$ & \cite{1986AJ.....91..602D} \\
&4.9  & 1983 May 20 & $0.26 \pm 0.08$ & \cite{1986AJ.....91..602D} \\
&4.9  & 1983 Dec 26 & $\le 0.18(3\sigma)$ & \cite{1986AJ.....91..602D} \\
&4.9  & 1984 Mar 17 & $0.24 \pm 0.05$ & \cite{1986AJ.....91..602D} \\
&15.0 & 1984 Nov 6 & $0.68 \pm 0.09$ & \cite{1986AJ.....91..602D} \\
&22.5  & 1999 Jan 06  & $ 1.7 \pm 0.2$ & \cite{2011AA...533A.107D} \\
&43.3  & 1999 Jan 06 & $ 3.3 \pm 0.4$ & \cite{2011AA...533A.107D} \\
&43.3  & 2004 Jan 25 & $ 3.34 \pm 0.08$ & \cite{2011AA...533A.107D} \\
&86.0  & 1985 Nov  & $ 21.4 \pm 7.5$ & \cite{1986AA...164..227A} \\
&108.4  & 1997 Nov - 2000 Jun & $ 20.09 \pm 0.69$ & \cite{2005AJ....129.2836C} \\
&217.8 & 1997 Nov - 2000 Jun  & $ 83.5 \pm 1.71$ & \cite{2005AJ....129.2836C} \\
&250.0  & 1986 Dec - 1989 Mar  & $ 78 \pm 8$ & \cite{1994AA...281..161A} \\
\hline
\rule{0pt}{3ex}    $\alpha$ Tau (K5 III)	&4.9  & 1983 Jan 21 & $\le 0.27(3\sigma)$ & \cite{1986AJ.....91..602D} \\
&4.9  & 1984 Nov 6 & $\le 0.22(3\sigma)$ & \cite{1986AJ.....91..602D} \\
&5.0  & 1997 Sep 27 & $\le 0.07(3\sigma)$		& \cite{2007ApJ...655..946W} \\
&8.5  & 1997 Sep 27 & $0.28 \pm 0.03$ 		& \cite{2007ApJ...655..946W} \\
&14.9 & 1997 Sep 27 & $0.95 \pm 0.08$ 		& \cite{2007ApJ...655..946W} \\
&15.0 & 1984 Nov 6 & $0.60 \pm 0.1$ 		& \cite{1986AJ.....91..602D} \\
&108.4  & 1997 Nov - 2000 Dec & $ 13.97 \pm 1.46$ & \cite{2005AJ....129.2836C} \\
&217.8 & 1999 Sep - 2000 Dec  & $ 25.78 \pm 5.64$ & \cite{2005AJ....129.2836C} \\
&250.0  & 1986 Dec - 1987 Jan & $ 51 \pm 6$ & \cite{1994AA...281..161A} 
\enddata
\label{tab:tab4}
\end{deluxetable*}

\section{DISCUSSION}
\subsection{Results Versus Previous Observations} \label{disc1}
Prior to and during the early operation of the `old' VLA, a small number of single dish radio observations reported the detection of flares from single red giants (e.g., \citealt{1981ApJ...245L..71B}). These transient radio events have never been re-observed however, even with more sensitive interferometers, suggesting that such detections were false (e.g., \citealt{1992MNRAS.254....1B}). The first definitive detection of thermal free-free emission from a luminosity class III single red giant at centimeter wavelengths was of $\alpha$ Boo at 6 cm \citep{1983ApJ...274L..77D,1986AJ.....91..602D}. Since then there has been a modest number of sporadic centimeter and millimeter observations of this star. In Table \ref{tab:tab4} we list the majority of these observations and plot their flux densities as a function of frequency in Figure \ref{fig:fig1}. In comparison to other single red giants, $\alpha$ Boo had been relatively well observed at radio wavelengths before this study, including detections in four VLA bands (i.e., Q, K, Ku, and C). No Ku band receivers were available during the commissioning phase of the VLA in early 2011 so we can compare three of our detections with previous ones.

Previous detections of $\alpha$ Boo at 6 cm ranged from a 3$\sigma$ upper limit of 0.18 mJy to a 3$\sigma$ detection at 0.39 mJy. Our 6 cm value agrees to within $\sim$10$\%$ of the highest S/N (5$\sigma$) value of \cite{1986AJ.....91..602D}. There is no significant difference between our 1.3 cm value and that of \cite{2011AA...533A.107D}. There is however a notable difference in flux density values at 0.7 cm  where \cite{2011AA...533A.107D} report values that are lower than ours by over 40\%. Such a level of chromospheric variability seems rather high and would be unexpected from such supposedly inactive stars. Another possibility for the difference in values is that the longer cycle time used by \cite{2011AA...533A.107D}, which was over double our value, may cause larger phase errors and thus lower final flux density values. Future high frequency VLA observations of $\alpha$ Boo will clarify this discrepancy at 0.7 cm but past detections at longer wavelengths appear to be in good agreement with our data.
